---
title: "tidytuesday_two"
format: html
editor: visual
---


### Loading in necesseary libraries

```{r message=FALSE, results='hide'}
library(janitor)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)
```


## Loading the data

```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 7)
tuesdata
historical_spending <- tuesdata$historical_spending
gifts_age <- tuesdata$gifts_age
gifts_gender <- tuesdata$gifts_gender

```


### Adding a column displaying the changes for different variables

```{r}
per_person <- historical_spending$PerPerson
PerPerson <- historical_spending$PerPerson
for(i in seq_along(PerPerson)){
  if(i==1){per_person[i] <- 0}else{per_person[i] <- PerPerson[i] - PerPerson[i-1]}
}

historical_spending <- historical_spending %>% 
  mutate(change_in_spending = per_person)


```


- The data was collected by the National Retail Federation in the US, by surveying customers on how they planned on celebrating valentine's day. The data I will be using has responses on different amount spent on different gifts or celebrations from 2010 to 2022.

/
/

### How does the percent of people celebrating valentine's and average spending on celebration change with years?

- The overall percent decrease in number of people celebrating across the whole 12 yr period is approximately 8 percent.
- The average amount of money spent by each person has increased by approximately $100 over the years, and this might be partly due to inflation. 
- The largest and most sudden drop in percent of people celebrating was from 2013 to 2014, and in this time period, we see that the average amount of spending by each person has the slightest increase.


<details>
<summary>View Code</summary>

```{r}
#changes in number of people celebrating over the years
historical_spending %>%
  ggplot(aes(x = Year, y = PercentCelebrating)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 2010:2022) +
  labs(title = "Percentage of people celebrating valentine's day over the years",
       x = "Year",
       y = "Percentage of people celebrating")

# average spending over the years
historical_spending %>%
ggplot(aes(x = Year, y = PerPerson)) +
  geom_segment(aes(xend = c(tail(Year, n = -1), NA), 
                   yend = c(tail(PerPerson, n = -1), NA)),
               arrow = arrow(length = unit(0.4, "cm")),
               color = 2) +
  geom_point(size = 2, color = 2) +
  labs(title = "Changes in average amount spent by each person celebrating Valatines over the years",
       x = "Year",
       y = "Average amount spent by each person")
```


</details>



### 

<details>
<summary>View Code</summary>

```{r}

historical_spending %>% 
  ggplot(aes(x = Year)) +
  geom_path(aes(y = Candy, color = "Candy"), linewidth = 0.9) +
  geom_line(aes(y = Flowers, color = "Flowers"), linewidth = 0.9) +
  geom_line(aes(y = Clothing, color = "Clothing"), linewidth = 0.9) +
  geom_line(aes(y = GiftCards, color = "Gift cards"), linewidth = 0.9) +
  geom_vline(aes(xintercept = Year), linetype = 3, color = "darkgrey") +
  scale_x_continuous(breaks = 2010:2022) +
  scale_y_continuous(breaks = seq(8,24,by=2)) +
  labs(y = "Amount spent", title = "Changes in amount spent on different items for valentine's celebration") +
  geom_text(aes(label = floor(change_in_spending), x = Year - 0.5, y = Candy+4, color = "Change in avr amt spent"), color = "darkgrey", size = 3) +
  theme_classic()

```

</details>





